---
title: "Merge Author Data"
author: "Ben Schmidt"
date: "2019-02-19"
output: html_document
---


```{r global_options, include=FALSE}
library(knitr)
library(magrittr)
library(ggplot2)
library(dplyr)

#opts_chunk$set(eval=FALSE, warning=FALSE, message=FALSE)
```


# Joins

How do you merge in classification data to our set? This requires an additional abstraction about working with tabular data: the **join**.

The classical case for a join is when you 

I've added a file to the server that contains labels for Library of Congress classifications.
This has a number of different columns: the code you've already seen, a description 
of the subject for the code, and a more general column that I've defined somewhat capriciously.

```{r}
library(tidyverse)
classifications = read_tsv("../data/LOC_classification.tsv")
classifications
```

This should work well with the LOC gender data; here's an example using it with a database of book titles you should have in your local folder.

``` {r}
shortTitles = read_tsv("../data/booktitlesShort.tsv")

head(shortTitles)
```

This can be done with the `inner_join` function of dplyr. By default it merges based on shared column names; you can feed it a precise list of columns to match on, but it is often easier to simply use `mutate` to rename the columns.

``` {r}

labelled = classifications %>% mutate(lc1 = class) %>% inner_join(shortTitles)

```

`inner_join` drops rows that it ca . Only about a third of the books in the `shortTitles` folder have an lc1 that matches the classifications field. The `nrow` function tells how many rows a datafraem has: here you can see that joining it to the genre lookup loses almost two-thirds the entries.

``` {r}
nrow(shortTitles)
nrow(labelled)

```

To see what's being dropped, you can use the `anti_join` function that shows only elements in the first frame that are unmatched in the second one. Here you see the other way of doing a join: adding a 'by' argument to the join function.

``` {r}
shortTitles %>% anti_join(classifications, by = c("lc1" = "class")) %>% sample_n(30) %>% select(lc1)
```

If you want to keep all elements from the left frame and just put `NA` values into the various fields defined by the right one, you can use the `left_join` function which prioritizes things from the left. (`outer_join` privileges both fields, and `right_join` privileges the second one.) "Left" and "right" refer, obviously, to the order that you pass them t the function.


``` {r}
counts = classifications %>% mutate(lc1 = class) %>% inner_join(shortTitles) %>% 
  mutate(workyear = as.numeric(workyear)) %>%
  group_by(BenSubject, workyear) %>% 
  summarize(count=n())

counts %>%
  ggplot() + geom_bar(stat="identity") + aes(x=workyear, y = count, fill=BenSubject) + lims(x = c(1850, 1880))
```
