---
title: "11 Geocomputation"
author: "Ben Schmidt"
date: "March 28, 2019"
output: html_document
---

You may have looked in other classes at geodata in a framework like QGIS.

In R, you can deal with geodata directly by reading it in using the `sf` package.

(There are other, older ways to read in data; I do not recommend them.)

```{r setup, include=FALSE}
library(maps)
library(tidyverse)
library(sf)
library(leaflet)
# install.packages("sf")
```
An SF object is basically a data frame with a special `geometry` field that indicates what something's shape is, and some additional data about projections.

There are two different ways to make an sf object. One is to just read in any shapefile, geojson, or other object.

```{r}
states <- st_read("../data/maps/cb_2017_us_state_5m.shp")
states
```

Another is to *convert* an existing object by telling where the latitude and longitude fields are. 

We can do this with our city data from earlier in the semester.

```{r}
cities <- read_csv("../data/CESTACityData.csv") %>% filter(!is.na(LAT)) %>% st_as_sf(coords = c("LON", "LAT"), crs = 4326)
```

```{r}
cities %>%
  st_crop()
filter(LAT > 10)
sample_n(100)
library(leaflet)

ne <- c("ME", "MA", "VT", "RI", "CT", "NH")
ne_top_pops <- cities %>%
  filter(ST %in% ne) %>%
  gather(year, pop, `1790`:`2010`) %>%
  group_by(CityST) %>%
  arrange(-pop) %>%
  slice(1) %>%
  mutate(year = as.numeric(year))
```

To build a map, you just composite layers in ggplot. Choose attractive colors and sensible geometries.

```{r}

# %>% st_transform(crs=5070)

ggplot(ne_top_pops) +
  geom_sf(data=states, fill='beige', lwd=0) + 
  geom_sf(data = states %>% filter(STUSPS %in% ne)) + 
  geom_sf(aes(color = `year`, size = `pop`)) +
  theme_void() +
  coord_sf(xlim = c(-79, -67), ylim = c(40, 48)) +
  scale_color_viridis_c() + 
  theme(panel.background = element_rect(fill = "aliceblue"))

```


```{r cars}
leaflet(ne_top_pops) %>% addCircleMarkers(color = ~year, radius = ~ .03 * sqrt(pop)) %>% addTiles()
```

```{r}

```

```{r pressure}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
